---
title: "Assignment 6"
author: "Marissa Dyck"
date: "Due Tuesday March 25th @11:59pm"
output: 
 html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Assignment 4: Data visualization

library(tidyverse)
library(cowplot)
library(leaflet)
library(ggpubr)
library(RColorBrewer)
library(stringr)
```

## Submission
Please email me a copy of your R script for this assignment by end of day **Friday March 29th** with your first and last name followed by assignment 3 as the file name (e.g. 'marissa_dyck_assignment4.R')

> You should always be following best coding practices (see Intro to R module 1) but especially for assingment submissions. Please make sure each problem has its own header so that I can easily navigate to your answers and that your code is well organized with spaces as described in the best coding practices section and comments as needed. 

## 1 Boxplot

Using the built in *PlantGrowth* dataset make a plot that meets the following requirements

1. Boxplot of weight for each group

2. Filled by group (choose some greyscale colors) and outlined in black

3. Informative axis titles and text

4. Readable text size

5. No gridlines

5. No legend

```{r eecho=TRUE, class.source = 'fold-hide'}

plant_boxplot <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  
  # add boxplots filled by group 
  geom_boxplot(aes(fill = group),
               color = 'black') + # outline in black
  
  labs(x = 'Treatment group',
       y = 'Weight (g)') + # I am just guessing it is in grams since the R documentation did not say
  
  # select greyscale colors
  scale_fill_manual(values = c('#DDDDDD', '#888888', '#444444')) +
  
  # change axis labels
  scale_x_discrete(labels = c('Control', 'Treatment 1', 'Treatment 2')) +
  
  # select theme with no gridlines
  theme_classic() +
  
  # remove legend and increase text size
  theme(legend.position = 'NONE',
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) 
```

## Scatterplot

Using the built in *trees* data make a plot that meets the following requirements

1. Scatterplot with girth as a function of height

2. Change the points plus signs

3. Informative axis titles and text

4. Readable text size

5. Informative title that is centered at the top

6. axis with breaks every 5 units starting at a number divisible by 5

7. Gridlines
```{r echo=TRUE, class.source = 'fold-hide'}

# specify data and variables
trees_scatterplot <- ggplot(trees, aes(x = Height, y = Girth)) +
  
  # add points as plus signs
  geom_point(shape = 3,
             size = 4) +
  
  # create informative labels and title
  labs(x = 'Height (ft)',
       y = 'Diameter (in)',
       title = 'Black Cherry Tree Measurements') +
  
  # add breaks based on range of data
  scale_x_continuous(breaks = seq(60, 85, by = 5)) +
  scale_y_continuous(breaks = seq(5, 20, by = 5)) +
  
  # choose theme with gridlines
  theme_bw() +
  
  # adjust text and title
  theme(plot.title = element_text(hjust = 0.5,
                                  size = 16),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
```

## 3 Read/format data

> For the next three problems you will be using the 'bear_2008_2016.csv' data


Please read in the data according to the following guidelines

1. Save it as 'bears'

2. Set all variable names to lowercase

3. Keep only points where damage is 1

4. Keep only the following columns (damage, year, targetspp, bear_abund, landcover_code, altitude, human_population, disto_to_forest, and dist_to_town)

*You may have to do some other data manipulation later on for each graph
```{r echo=TRUE, class.source = 'fold-hide'}

# read in bear data

bears <- read.csv('data/raw/bear_2008_2016.csv') %>% 
  
   # set column names to lowercase
  set_names(
    names(.) %>%  # the period here is a placeholder for the data it tells R to use the element before the last pipe (e.g., turtles.df)
      tolower()) %>% 
  
     # select only damage points
  filter(damage == 1) %>% 
  
  # select specified columns
  select(damage, year, targetspp, bear_abund:dist_to_town)

str(bears)
  
  
# filter data to only damage points

```

## 4 Bar graph

Create a plot using the bear data so it meets the following criteria

1. Bar graph showing the total number of livestock events per year per livestock type

2. Bars colored by livestock type (manually set the colors)

3. Informative title, caption, and axis with readable text size

4. y axis breaks every 20

5. No ticks on x - axis and years at a 45 degree angle

6. no legend

7. No gridlines and only borders on the left and bottom of graph

```{r echo=TRUE, class.source = 'fold-hide'}

# mutate data so year and targetspp are factors

bears_summary <- bears %>% 

  
  # change variables to factor
  mutate(year = as.factor(year),
         targetspp = as.factor(targetspp))


bear_damage_bargraph <-  ggplot(bears_summary, aes( x = year)) +
  
  # add bars
  geom_bar(aes(fill = targetspp),
           position = position_dodge()) +
  
  # specify colors and labels
  scale_fill_manual(name = 'Livestock type',
                    values = c('#A31E17', '#4F3124', '#AF9985'),
                    labels = c('Other', 'Cows', 'Sheep')) +
  
  scale_y_continuous(expand = c(0,0),
                     breaks = seq(0, 80, by = 20)) +
  
  # add labels, caption, and title
  labs(x = 'Year',
       y = 'Number of events',
       title = 'Reported livestock damage by brown bears',
       caption = str_wrap('The total number of reported livestock damage events caused by brown bears in Romania per year and livestock type (tan = sheep, brown = cows, red = other livesotock [e.g. chickens]) from 2008 to 2016.',
                          width = 90)) +
  
  
  theme_classic() +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_text(size = 12,
                                   angle = 45,
                                  vjust = 0.7),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18),
        legend.position = 'NONE',
        plot.caption = element_text(size = 12,
                                    hjust = 0))
```

## 5 Combine multiple plots

For this problem you'll need to make a change to the data

* Create a new column the combines values from the landcover code    
  art_surfaces = 112, 121, 131    
  ag = 211, 221, 222, 242, 243    
  open = 231, 321   
  forest = 311, 312, 313, 324   

Then you will need to make **two** graphs using this new column


1. Grouped landcover type and dist_to_forest

2. Grouped landcover type and dist_to_town

Then **combine** them into one plot with a common legend.


The graphs should all meet the following criteria

1. Raw data plotted *hint: use `geom_jitter()` to avoid having all points stacked on top of eachother*

2. Points shaped **and** colored by targetspp

3. manually set the color *hint to get one legend per plot you will also need to manually set the shape using the same name and labels*

4. Informative and readable axis and text
  

```{r echo=TRUE, class.source = 'fold-hide'}

# create vectors for landcover types

art_surfaces <- c(112,121,131) 

ag <- c(211,221,222, 242,243) 

open <- c(231, 321) 

forest <- c(311,312,313, 324) 


bears_cows <- bears %>% 
  
  # keep only data for cows
  #filter(targetspp == 'bovine') %>% 
  
  # combine landcover types using case_when
  mutate(landcover_grouped = case_when(landcover_code %in% art_surfaces ~ 'art_surfaces',
                                       landcover_code %in% ag ~ 'ag',
                                       landcover_code %in% open ~ 'open',
                                       landcover_code %in% forest ~ 'forest'))

str(bears_cows)

# create vectors for plot colors and labels

spp_colors <- c('#A31E17', '#4F3124', '#AF9985')

spp_shapes = c(17, 16, 15)

spp_labels <- c('Other', 'Cows', 'Sheep')


# plot 1

# specify data and variables
plot_5.1 <- ggplot(bears_cows, aes(x = landcover_grouped, y = dist_to_forest)) +
  
  # add points
  geom_jitter(aes(color = targetspp,
                  shape = targetspp),
              width = 0.4, # this changes the amount of 'jitter'
              alpha = 0.6, # this sets opacity
              size = 2) +
  
  # manually set colors
  scale_color_manual(name = 'Livestock type',
                     values = spp_colors,
                     labels = spp_labels) +
  scale_shape_manual(name = 'Livestock type',
                     values = spp_shapes,
                     labels = spp_labels) +
  scale_x_discrete(labels = c('Agriculture', 'Artifical', 'Forest', 'Open')) +
  
  # informative labels
  labs(x = 'Landcover type',
       y = 'Distance to forest (m)') +
  
  # add theme elements
  theme_bw() +
    theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))

# plot 2

# specify data and variables
plot_5.2 <- ggplot(bears_cows, aes(x = landcover_grouped, y = dist_to_town)) +
  
  # add points
  geom_jitter(aes(color = targetspp,
                  shape = targetspp),
              width = 0.4, # this changes the amount of 'jitter'
              alpha = 0.6, # this sets opacity
              size = 2) +
  
  # manually set colors
  scale_color_manual(name = 'Livestock type',
                     values = spp_colors,
                     labels = spp_labels) +
  scale_shape_manual(name = 'Livestock type',
                     values = spp_shapes,
                     labels = spp_labels) +
  scale_x_discrete(labels = c('Agriculture', 'Artifical', 'Forest', 'Open')) +
  
  # informative labels
  labs(x = 'Landcover type',
       y = 'Distance to town (m)') +
  
  # add theme elements
  theme_bw() +
    theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))

plot_combined <-  ggarrange(plot_5.1, plot_5.2,
          common.legend = TRUE)
```

# 'Answer Key'

## 1 Boxplot
```{r, echo = FALSE}

plant_boxplot

```

## 2 Scatterplot
```{r, echo = FALSE}

trees_scatterplot

```

## 3 Read/format data

```{r}

str(bears)

```


## 4 Bar graph
```{r, echo = FALSE}

bear_damage_bargraph
```

## 5 Combined plots
```{r, echo = FALSE}

plot_combined
```



