---
title: "Assignment 2"
author: "Working with data"
date: "Due TBD"
output: 
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)

library(tidyverse)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Assignment 2: working with data

library(tidyverse)
```

## Submission

Submit you R script as a .R (or .Rmd if using markdown) file to Brightspace.

> Please make sure your submission includes your name and the assignment number in the filename 


## Grading

> You should always be following best coding practices (see Intro to R module 1) but especially for assingment submissions.

To receive full credit for each assignment  

* Wherever possible prioritizxe the use of tidyverse functions  
* Please make sure each problem has its own header so that I can easily navigate to your answers 
* Ensure you have comments that explain what you are doing   
* Long code chunks should be broken up with spaces and comments to explain what is happening at each step    
* Object names should be lowercase and short but descriptive enough that they aren't confused with other objects *(For example data1 and data2 are not good names for dataframes you are working with)*    
* Just because your code runs doesn't mean it did what you think it did, always check your data/objects to ensure any functions were performed correctly (there are several ways to do this)    



## 1 Export data

Save the altered turtles data as a comma separated file to the *data/processed* folder in your working directory using the 'readr' package and name it 'turtles_tidy'
```{r echo=TRUE, class.source = 'fold-hide', message=FALSE}
# includes code for altering turtle data from the lab script

# read in turtle data from .txt file using tidyverse read_delim() function
turtles.df <- read_delim('data/raw/turtle_data.txt',
                         delim = '\t') %>% 
  
   # set column names to lowercase for easier coding later
  set_names(
    names(.) %>%  
      tolower()) %>% 
  
  # rename columns to shorter names for easier coding later, code reads (new_name = old_name)
  rename(tag = tag_number,
         c_length = carapace_length,
         h_width = head_width) %>% 
  
  # change the sex variable to type factor
  mutate(sex = as.factor(sex))
```

```{r echo=TRUE, class.source = 'fold-hide'}
# answer 1

# save the new clean data as a csv file using tidyverse write_csv() function. first provide the object name and then the file path (folder and subfolders), name (the file name), and extension (.csv) you want to save it as in quotes
write_csv(turtles.df,
          'data/processed/turtles_tidy.csv')

```


 

## 2 Import and rename data

> First if you haven't, download the [brown bear damage data (bear_2008_2016.csv)](data/raw/bear_2008_2016.csv) and save it in the data/raw folder.

* Import the brown bear dataset using the appropriate function from the 'readr' package (e.g. tidyverse) and save it to the environment as bear_data.    
* In the same code chunk set the column names to lowercase and    
* change dist_to_town to m_to_town and dist_to_forest to m_to_forest (m for meters)    
* Finally, use one of the functions we've covered to view/print your data to check that it worked.

```{r echo=TRUE, class.source = 'fold-hide'}
# read in bear data and assign to env. as bear_data
bear_data <- read_csv('data/raw/bear_2008_2016.csv') %>% 
  
  # set column names to lowercase
  rename_with(tolower) %>% 
  
  # rename columns to include units
  rename(m_to_town = dist_to_town,
         m_to_forest = dist_to_forest)

# view data to check changes
head(bear_data)


# OR

bear_data <- read_csv('data/raw/bear_2008_2016.csv') %>% 
  
  # set column names to lowercase using the example from class which was written for purrr functions (sorry forgot to update)
  set_names(
    names(.) %>%  
      tolower()) %>% 
  
  # rename columns to include units
  rename(m_to_town = dist_to_town,
         m_to_forest = dist_to_forest)
  

# view data
head(bear_data) 
```

## 3 Change variable type 

* Check the internal structure of the bear dataset    
* Change 'targetspp' from **character** to **factor** using *dplyr* make sure to overwrite the previous bear data so the changes are saved for your object    
* Use the 'levels()' **function** to check if this worked. HINT: Use the help file if you aren't familiar with 'levels()'

```{r echo=TRUE, class.source = 'fold-hide'}
# check how variables read in for the data 
str(bear_data)

# change targetspp to type factor with dplyr and mutate
bear_data <- bear_data %>% 
  
  # change variable to type factor
  mutate(targetspp = as.factor(targetspp))


# check levels of factor variable
levels(bear_data$targetspp)
```

## 4 Subsetting 

* Now that we know there are 3 livestock types (groups), subset the data to include only the rows for 'ovine' 
* In the same code chunk select the columns damage, year, month, bear_abund, landcover_code, and altitude
* Save this to your **environment** as 'bear_sheep_data'
```{r echo=TRUE, class.source = 'fold-hide'}

bear_sheep_data <- bear_data %>% 
  
  # return only rows for sheep
  filter(targetspp == 'ovine') %>% 
  
  # select specified columns
  select(damage:month, bear_abund:altitude) # most parsimonious way to do it but you could have also done


# select(damage, year, month, bear_abund, landcover_code, altitude)

# or 

# select(1:3, 7:9) I do not prefer this one as it's easy for someone reading your code to misinterpret and not change the column names if needed

head(bear_sheep_data)
```

## 5 Summarise 

* Using the bear_sheep_data and the `summarise()` **function**, calculate the mean, sd and SE (the formula for SE is sd / sqrt(n)) for altitude. *This last part might be tricky at first but give it a try and remember you can always google things and check the 'help' files; and if that fails 'phone a friend or me'*

```{r echo=TRUE, class.source = 'fold-hide'}

bear_sheep_data %>% 
  
  # use summarise to calculate some summary statistics from the data, mean, sd, and SE
  summarise(mean_alt = mean(altitude),
            sd_alt = sd(altitude),
            se_alt = sd_alt/sqrt(length(altitude))) 

# you could also look up the length for altitude in the console and type the number here directly but this is more flexible and readable

# or another possible answer

bear_sheep_data %>% 
  summarise(mean_alt = mean(altitude),
            sd_alt = sd(altitude),
            se_alt = sd_alt/sqrt(n()))

```
