---
title: "Assignment 1"
author: "Base R"
date: "Due Tuesday Jan 14th @11:59pm"
output: 
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)

library(tidyverse)
```


Below are some exercises for your first stats assignment based on the material we just covered. For all exercises the results will be available on this website after I've graded the assignments. These are meant to test your knowledge of the material we've covered and help you learn to work with real data. These are meant to mimic working in R in the real world and you will have to modify code from the module and possibly learn to use a new function to complete these assignment, just like in real life. Remember you can always look up information for a particular **function** using `?function name`to open the **help** window. 

## Submission

Brightspace/TBD   

## Grading

> You should always be following best coding practices (see Intro to R module 1) but especially for assingment submissions.

To receive full credit for each assignment   

* Please make sure each problem has its own header so that I can easily navigate to your answers 
* Ensure you have comments that explain what you are doing   
* Long code chunks should be broken up with spaces and comments to explain what is happening at each step    
* Object names should be lowercase and short but descriptive enough that they aren't confused with other objects *(For example data1 and data2 are not good names for dataframes you are working with)*    
* Just because your code runs doesn't mean it did what you think it did, always check your data/objects to ensure any functions were performed correctly (there are several ways to do this)    


## 1 Create a vector
Create a vector called 'myvec' using any of the methods you learned with numbers 1 to 10. *Note there are multiple ways to do this*. 

```{r echo=FALSE}
# (1 pt)

# answer 1 (the most parsimonious)
myvec <- 1:10

# answer 2
myvec <- c(1, 10)

# answer 3 (time consuming)
myvec <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

# answer 4 (unecessarily complicated for this exercise but works)
myvec <- seq(1, 10, by = 1)

myvec
```

## 2 Create a matrix with `rbind()`
Create a 3 row by 2 column matrix named 'mymat'. Use the `rbind()` **function** to bind the following three rows/vectors together:

```
c(1,4) 
c(2,5) 
c(3,6)
```

```{r echo=FALSE}
# (1 pt)

mymat <- rbind(
  c(1,4),
  c(2,5),
  c(3,6)
)

mymat
```

## 3 Extracting objects from data
Get the names of columns in the data frame you created earlier 'mydf'. *Hint see the R functions to explore data section*. Then extract all rows for column 5 by name, do the same thing using the element position e.g. `[]`

```{r echo=FALSE}
# code to create mydf
my.data <- data.frame(
  Obs.Id = 1:100,
  Treatment = rep(c("A","B","C","D","E"),
                  each = 20),
  Block = rep(1:20,
              times = 5),
  Germination = rpois(100,
                      lambda = rep(c(1,5,4,7,1),
                                   each = 20)),   # random poisson variable
  AvgHeight = rnorm(100,
                    mean = rep(c(10,30,31,25,35,7),
                               each = 20))
)

mydf = my.data[21:30, ]  

# answer for problem 1

# (3 pts)

names(mydf)

mydf$AvgHeight

mydf[ , 5]

```

## 4 Create a matrix
Create a new matrix called 'mymat2' that includes all the data from columns 3 to 5 of data frame mydf. *HINT: use the `as.matrix()` function to coerce a data frame into a matrix. Since we didn't cover this function you may need to look it up in the help files.*

> Note your values for some columns may be slighly different since the code to create mydf uses random number generators. 

```{r echo=FALSE}

# (1 pt)

mymat2 <- as.matrix(mydf[ , 3:5])

mymat2

```


## 5 Create a list
Create a list named 'mylist' that is composed of a    
  - vector: 1:3,     
  - a matrix: matrix(1:6, nrow = 3, ncol = 2),      
  - and a data frame: data.frame(x =c (1, 2, 3), y = c(TRUE, FALSE, TRUE), z = c("a", "a", "b")).
  
```{r echo=FALSE}

#create an empty list
mylist <- list()

# add a vector of 1 to 3 to the list
mylist[[1]] <- 1:3

# add a matrix to the list
mylist[[2]] <- matrix(1:6,
                      nrow = 3,
                      ncol = 2)

# add a data frame to the list
mylist[[3]] <- data.frame(x = c(1, 2,3 ),
                          y = c(TRUE, FALSE, TRUE),
                          z = c("a", "a", "b"))

# print the list
mylist

```

## 6 Extracting objects from lists
Extract the second and third observation from the 1st column of the data frame in 'mylist' (the list created above). 

```{r echo=FALSE}
# multiple ways to do this

# answer 1 - reference the element in the list [[3]] then the rows and columns you want from that element [2:3, 1]
mylist[[3]][2:3, 1]

# answer 2 call mylist then reference the position of the data frame in the list [[3]], then the column in the data frame [[1]], and finally the observations within that columns c(2, 3)
mylist[[3]][[1]][c(2, 3)]

```


