---
title: "Welcome to the R Crash Course!"
date: "Spring 2025"
output:
  html_document:
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 * **When?** Tuesday's @2:30 - 4:20pm & Friday's @2:30 - 3:20pm       
 * **Where?** Clearihue A312        
 * **What to prepare?** Please bring your laptop computer! Also, install R and RStudio prior to the course and following the instructions below on 'To do before lab 1'.    
 * **Office hours:** My "office hours" will immediately follow lab each week for 1 hour in Clearihue A312 (assuming there is no class after us). I encourage you to stay and work on the practice problems with your peers and I can answer any questions you have regarding the lab material, assignments, working with your own data etc. during this time.   
 
## Facilitators
[Marissa Dyck](https://github.com/marissadyck)  
University of Victoria    
School of Environmental Studies     
Email: [marissadyck@uvic.ca](mailto:marissadyck@uvic.ca)   

[Jason Fisher](http://www.acmelab.ca/the-lab.html)  
University of Victoria    
School of Environmental Studies   
Email: [fisherj@uvic.ca](mailto:fisherj@uvic.ca)   

## Contributors
This workshop is based on content from an R Bootcamp run by Dr. Kevin Shoemaker at University of Nevada Reno (https://github.com/kevintshoemaker/R-Bootcamp)

[Kevin Shoemaker](http://naes.unr.edu/shoemaker/)	
University of Nevada Reno  
Department of Natural Resources and Environmental Science       
Email: [kshoemaker@cabnr.unr.edu](mailto:kshoemaker@cabnr.unr.edu)

## About

The statistical programming software 'R' is one of the fundamental tools for modern data exploration and is a useful tool for data processing, statistical analysis,  and production of high-quality figures. 

This workshop is designed for beginner to intermediate R users and will focus on increasing comfort and familiarity with using R and RStudio, R syntax, and troubleshooting code. We will begin with the basics (what is R? How to install R and RStudio? Navigating and understanding the layout of Rstudio) and continue onto data manipulation and formatting, visualizing data, working with packages, and troubleshooting code. The main goal of the workshop is to ensure participants have enough proficiency and confidence with data operations and programming in R to engage in productive, self-directed learning and problem-solving.  

All code will be available as scripts that you can download from this website (at the top of each module page on this website) and load up in RStudio. That way you won't need to constantly copy and paste from the web!  

## Course Schedule

See Brightspace for most up-to-date schedule        

## To do before lab 1

Please complete the following tasks BEFORE the first lab on Tuesday January 14th. We have a lot of material to cover in a short time and completing the following tasks will save time at the start of lab and ensure everyone is ready.

### Install R and RStudio

Before we dig in and get started with the modules, you should install R and RStudio. Even if you have installed R and RStudio before you should ensure you have the latest version, if you aren't sure how to check use the links to install the programs again just to be sure and we will cover how to check your version during the workshop. Here are some links to help you get started:

[Download and install R](https://cran.r-project.org/)        
[Download and install RStudio](https://www.rstudio.com/products/rstudio/download/)    (use free version!)      


### Download cheat sheets
Also, it can be very helpful to print out R 'cheat sheets' and bring that with you or save them as PDFs to your hard drive. We will try to have some physical copies available during lab. Here are some links:

[Base R cheatsheet](https://iqss.github.io/dss-workshops/R/Rintro/base-r-cheat-sheet.pdf)           
[R reference card](https://cran.r-project.org/doc/contrib/Short-refcard.pdf)  
[Various R cheatsheets](https://posit.co/resources/cheatsheets/?type=posit-cheatsheets&_page=2/) (I recommend the following 3: 'Data tidying with tidyr', 'Data transformation with dpylr', 'Apply functions with purrr', and 'Data visualization with ggplot2' cheat sheets for this course)

### Download data

To ensure we maximize time in lab please also download the data files for this course from the [Data](data.html) tab and save this in a folder on your hard drive. *We will move these to a specific folder for the course when we cover folder organization on the first day of lab*

Right click (or command-click for MAcs) each file. The .txt files will open in a viewer window and you will need to right click to save them. The .csv files should download directly in your downloads folder and you can move them where you want to save them. **Important DO NOT change the file names**.


## Go to first lab

Okay, now we're ready to go!


[--go to first lab module](mod_1_R101.html)




```{r echo=FALSE}
rmd2rscript <- function(infile="mod_5_purrr.Rmd"){    # function for converting markdown to scripts
  outfile1 <- gsub(".Rmd",".R",infile)
  outfile2 <- gsub(".Rmd",".txt",infile)
  close( file( outfile1, open="w" ) )   # clear output file
  close( file( outfile2, open="w" ) )   # clear output file
  con1 <- file(infile,open="r")
  con2 <- file(outfile1,"w")
  con3 <- file(outfile2,"w")
  stringToFind <- "```{r*"
  isrblock <- FALSE
  count=0
  
  while(length(input <- readLines(con1, n=1)) > 0){   # while there are still lines to be read
    isrblock <- grepl(input, pattern = stringToFind, perl = TRUE)   # is it the start of an R block?
    if(isrblock){
      while(!grepl(newline<-readLines(con1, n=1),pattern="```",perl=TRUE)){
        if(count>1){
          write(newline,file=con2,append=TRUE)
          write(newline,file=con3,append=TRUE)
        }
        count=count+1
      }
      isrblock=FALSE
    }
  }
  closeAllConnections()
}

rmd2rscript2 <- function(infile="LECTURE2.Rmd"){    # function for converting markdown to scripts
  outfile <- gsub(".Rmd",".R",infile)
  close( file( outfile, open="w" ) )   # clear output file
  con1 <- file(infile,open="r")
  con2 <- file(outfile,"w")
  stringToFind <- "```{r*"
  stringToFind2 <- "echo"
  isrblock <- FALSE
  #count=0
  blocknum=0
  
  while(length(input <- readLines(con1, n=1)) > 0){   # while there are still lines to be read
    isrblock <- grepl(input, pattern = stringToFind, perl = TRUE)   # is it the start of an R block?
    showit <- !grepl(input, pattern = stringToFind2, perl = TRUE)   # is it hidden (echo=FALSE)
    if(isrblock){
      blocknum=blocknum+1
      while(!grepl(newline<-readLines(con1, n=1),pattern="```",perl=TRUE)){
        if((blocknum>1)&((showit)|(blocknum==2))) write(newline,file=con2,append=TRUE)
        #count=count+1
      }
      isrblock=FALSE
    }
  }
  closeAllConnections()
}

```


```{r echo=FALSE, eval=FALSE}

rmd2rscript2("mod_2_baseR.Rmd")
rmd2rscript2("mod_3_data.Rmd")
rmd2rscript2("mod_4_data_manipulation.Rmd")
rmd2rscript2("mod_1_plots_baseR.Rmd")
rmd2rscript2("mod_2_ggplot.Rmd")

rmd2rscript2("mod_2_glm.Rmd")
rmd2rscript2("mod_5_purrr.Rmd")

## TODO: make sure there are pdf copies of all the modules... 

knitr::opts_chunk$set(echo = TRUE, message = FALSE)
rmarkdown::render("home.Rmd",rmarkdown::pdf_document())
rmarkdown::render("Intro_to_R_part1.Rmd",rmarkdown::pdf_document())
rmarkdown::render("day1_2.Rmd",rmarkdown::pdf_document())
rmarkdown::render("day1_3.Rmd",rmarkdown::pdf_document())

rmarkdown::render("day2_1.Rmd",rmarkdown::pdf_document())
rmarkdown::render("day2_2.Rmd",rmarkdown::pdf_document())

```




