---
title: "Data Visualization"
author: "Module 4: Beyond ggplot"
date: "2024"
output: 
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Data Visualization
# Module 4: Beyond ggplot


```
## Beyond ggplot
  
ggplot contains an incredibly rich and powerful set of tools for visualizing data. But ggplot and base R are not your only options! A variety of packages exist for visualization, including:
  
### ggplot extensions
  
  Several packages have been created that build off of ggplot's syntax with additional functions. You can find a list of them [here](https://exts.ggplot2.tidyverse.org/gallery/). 

### Technique-specific plotting libraries

Several analytical packages come with their own plotting functions that produce some very nice visualizations. There are dozens out there, but a few of them are [visreg](https://cran.r-project.org/web/packages/visreg/visreg.pdf) for regression plots, [corrplot](https://cran.r-project.org/web/packages/corrplot/index.html) (and its ggplot counterpart, [ggcorrplot](http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2)) for graphical presentation of correlation matrices, and [rpart.plot](https://cran.r-project.org/web/packages/rpart.plot/index.html) as a companion to the decision tree package rpart.

### Interactive plots: leaflet and plotly

Increasingly, scientific journals are providing platforms for interactive graphics on their websites to accompany published articles. Interactive plots are also popular for personal, lab, and organizational websites, and they can provide another option for your own data exploration. Two of the most popular in R are [plotly](https://plot.ly/r/), which offers a huge variety of 2D and 3D plots, and [leaflet](https://rstudio.github.io/leaflet/), which is specifically for mapping.


