---
title: "Untitled"
author: "Marissa Dyck"
date: "2024-01-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First we will look at how Purrr is useful for split-apply-combine operations. This is when we want to split (filter) a data set into parts such as by species, site, etc. apply a function to each part such as calculating a mean, median, etc. and then combine the resulting parts back together. For this we will use the built-in *iris* data set which contains measurements for three species of iris' 

First the long way, which you should be able to do after the last modules
```{r}
# first let's 'read in' the iris data and make a few simple modifications to make it easier to use

# create new object called iris_tbl (iris tibble)
iris_tbl <-  iris %>% 
  
  # make a tibble
  as_tibble(.) %>% 
  
  # set names to lowercase
  set_names(
    names(.) %>% 
      tolower())

```

To do this normally we'd have to write a function to calculate the what we want that can be applied to each part (species) of the full data set
```{r}
# Write a custom function that will calculate the mean sepal length of any of the iris species:
mean_sepal_length <- 
  function(x) {
    iris_tbl %>% 
      filter(species == x) %>% 
      pull(sepal.length) %>% 
      mean()
  }

# Use your custom function to generate a list of mean sepal lengths for the three iris species:

tibble(
  species = 
    iris_tbl %>% 
    pull(species) %>%
    unique(),
  sepal_length = 
    c(mean_sepal_length('setosa'),
      mean_sepal_length('versicolor'),
      mean_sepal_length('virginica')))

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
